\section{Dataset Overview}

\subsection{Data Source}

This analysis uses the RSITMD (Remote Sensing Image-Text Matching Dataset) \cite{rsitmd}, introduced by Yuan et al. in IEEE TGRS 2021. RSITMD is designed for cross-modal retrieval in aerial imagery and pairs satellite images with human-annotated English captions.

\textbf{Download:} \url{https://drive.google.com/file/d/1NJY86TAAUd8BVs7hyteImv8I2_Lh95W6/view} (Google Drive) | \url{https://pan.baidu.com/s/1gDj38mzUL-LmQX32PYxr0Q} (Baidu, password: NIST). Distribution includes GeoTIFF images (256$\times$256 px) and a JSON annotation file.

\subsection{Dataset Structure}

\begin{table}[h]
\centering
\caption{RSITMD Dataset Overview}
\label{tab:dataset_overview}
\begin{tabular}{|l|c|c|}
\hline
\textbf{Metric} & \textbf{Train} & \textbf{Test} \\
\hline
Total Images & 4,291 & 452 \\
\hline
Total Captions & 21,455 & 2,260 \\
\hline
Captions / Image & 5 & 5 \\
\hline
Land Use Categories & 33 & 32 \\
\hline
Avg Words / Caption & 10.26 & 10.20 \\
\hline
Vocabulary Size (raw) & 3,567 & 1,218 \\
\hline
Vocabulary Size (clean) & 3,470 & 1,143 \\
\hline
\end{tabular}
\end{table}

Each image has exactly five human-annotated English captions describing the same scene with different word choices (paraphrase-style). No public validation set is provided; practitioners split the training set internally. The compact vocabulary (3,470 words) and consistent caption length ($\mu\approx10$) reflect a standardized annotation guideline.

\subsection{Land Use Categories}

33 categories across five thematic clusters:

\begin{itemize}
    \item \textbf{Transportation}: airport, bridge, parking, port, railwaystation, viaduct
    \item \textbf{Residential}: denseresidential, mediumresidential, sparseresidential
    \item \textbf{Commercial \& Public}: center, church, commercial, school, stadium, playground
    \item \textbf{Natural Landscape}: bareland, beach, desert, farmland, forest, meadow, mountain, pond, river
    \item \textbf{Urban Features}: baseballfield, boat, industrial, intersection, park, plane, resort, square, storagetanks
\end{itemize}

Note: the \textit{plane} category is absent from the test split (32 categories instead of 33), affecting evaluation fairness for that class.

\subsection{EDA Pipeline Overview}

The EDA follows three parallel analysis pipelines:

\begin{enumerate}
    \item \textbf{Text Analysis.} Captions are tokenized and analyzed for word frequency, bigrams, vocabulary size, and caption-length distributions. Color words and spatial prepositions are extracted to quantify the visual vocabulary.

    \item \textbf{Image Analysis.} Two levels: (a) metadata from filenames --- category distribution and class imbalance ratio; (b) pixel-level --- per-category brightness, RGB channel ratios, and texture (intra-block pixel std) computed from raw image files.

    \item \textbf{Multimodal Joint Analysis.} Cross-modal alignment: caption variability across the five annotations per image, color word coverage, spatial relationship frequency, and cross-referencing between text and image visual profiles.
\end{enumerate}

\begin{figure}[H]
\centering
\resizebox{0.85\textwidth}{!}{
\begin{tikzpicture}[
    node distance=2.2cm,
    block/.style={rounded corners, minimum height=1.1cm, text width=3cm, align=center, font=\small},
    dataset/.style={block, fill=accent-light},
    modality/.style={block},
    pipeline/.style={block, fill=blue!15},
    output/.style={block, fill=orange!15},
]

% Row 1: input
\node[dataset] (dataset) {RSITMD Dataset\\{\tiny JSON + GeoTIFF}};

% Row 2: modalities
\node[modality, fill=blue!20, below=1.2cm of dataset] (text) {Text\\Captions};
\node[modality, fill=green!20, below=1.2cm of dataset] (img) {Image\\Pixels};

% Row 3: pipelines
\node[pipeline, right=3cm of text] (ta) {Text Analysis\\\emph{word freq, vocab, length}};
\node[pipeline, right=3cm of img] (ia) {Image Analysis\\\emph{brightness, texture, RGB}};

% Row 4: joint
\node[pipeline, right=2.5cm of ta] (mma) {Multimodal Analysis\\\emph{color words, spatial, variability}};

% Row 5: outputs
\node[output, right=2.5cm of mma] (figs) {Figures\\\emph{8 charts, 4 tables}};
\node[output, right=2.5cm of ia] (stats) {Statistics\\\emph{10+ metrics, per-cat}};

% Arrows
\draw[->, thick] (dataset) -- (text);
\draw[->, thick] (dataset) -- (img);
\draw[->, thick] (text) -- (ta);
\draw[->, thick] (img) -- (ia);
\draw[->, thick] (ta) -- (mma);
\draw[->, thick] (ia) -- (mma);
\draw[->, thick] (mma) -- (figs);
\draw[->, thick] (ia) -- (stats);
\draw[->, thick] (ta) -- (stats);

\end{tikzpicture}
}
\caption{EDA Pipeline: from raw RSITMD data to multimodal insights via three parallel tracks.}
\label{fig:pipeline}
\end{figure}
